---
title: "Project 4"
author: "Group 4"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r set up, include = F}
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(cache = T)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

library(tidyverse)
library(ggplot2)
library(anytime)
library(caret)
library(parallel)
library(krr)
library(dplyr)

cores <- detectCores()

source("../lib/cross_validation.R")
source("../lib/ALS_R_KRR.R")
source("../lib/ALSwithR1R3.R")
source("../lib/ALSwithP3.R")
```

# Introduction
In this project, our group implement matrix factorization by focusing on different versions of alternating least squares algorithm for different Regularizations and Kernel Ridge Regression as Post-processing. The objective is to produce a good prediction of users’ preferences for movies on on the MovieLens dataset. For this project, our team #4 is assigned with Pairing combination 12 + 14 from the Collaborative. For evaluation, we compared RMSE results for different methods. Our group used R language to product model and reports. The specific technique used for the combination are illustrated below. For more details for methods, please also see (https://github.com/TZstatsADS/ADS_Teaching/blob/master/Projects_StarterCodes/Project4-RecommenderSystem/doc/Proj4_pairings_2020_Spring.pdf).


# Models

`Alternating Least Squares + Kernel Ridge Regression`

`Alternating Least Squares + Penalty of magnitudes + Temporal Dynamics + Kernel Ridge Regression`


# Model 1: Alternating Least Squares + Kernel Ridge Regression
## Step 1. Load Data and Train-Test Split
In this step, we loaded the data and added variable timediff for temporal dynamics, which is the difference between the user's rating time and the average rating time of this user. Then we splitted the original dataset to training set(80%) and test set(20%)
```{r}
# Convert timestamp and set up time difference
data <- read.csv("../data/ml-latest-small/ratings.csv")
data$timestamp <- anydate(data$timestamp)
data_R <- data %>% 
  group_by(userId) %>% 
  mutate(timediff = (timestamp - mean(timestamp)) %>% 
           as.numeric) %>% 
  ungroup() %>%
  arrange(timestamp)

set.seed(1)

# train-test split (.8/.2)
train_ind <- createDataPartition(data$userId, p=.8, list=F)

data_train <- data_R[c(train_ind), ]
data_test <- data_R[-c(train_ind), ]

train <- data_train[, 1:3]
test <- data_test[, 1:3]

```

## Step 2. Model Setup
### 1. Equations recap
### Alternating Least Squares
$$min_{q^{\star}p^{\star}} \sum_{(u,i) \in K} (r_ui-q_i^Tp_u)^2 + \lambda(\sum _i n_{q_i} \| q_i\|^2 +\sum _i n_{p_u} \| p_u\|^2 )$$
ALS technique rotate between fixing the $q_i$'s and fixing the $p_u$'s. When all pu's are fixed, system recomputes the $q_i$'s by solving a least-squares problem, and vice versa. This ensures that each step decreases object function until convergence. 

$f$: dimension of latent factors

$q_i$: factors associated with item i , measures the extent to which items possesses those factors

$p_u$: factors associated with user u, measures the extent of interest that user has in an item  are high on corresponding factors.

### Post-processing
Kernel Ridge Regression:
$$\hat{\beta} = (X^{T}X + \lambda I)^{-1}X^{T}y $$
$$\hat{y_{i}} = {x_i}^T\hat{\beta}$$
Equilvalent to:
$$\hat{\beta} = X^T(XX^T + \lambda I)^{-1}y$$

$$\hat{y}_{i} = K(x_i^{T}, X)(K(X,X) + \lambda I)^{-1}y$$
$y$: vector of movies rated by user i
$X$: a matrix of observations - each row of X is normalized vector of features of one movie j rated by user i: 
$$x_{j_2} = \frac{q_j}{\|q_j\|}$$

$K$: 
$$K(x_{i}^T, X) = x_i^Tx_j$$ 
$$K(x_{i}^T, X) = \exp(2(x_i^Tx_j -1))$$

We first defined similarity between movies by the above formula. Then we used this similarity S to apply Kernel Ridge Regression prediction. 

### 2. Algorithm Setup
For matrix factorization, with the method in paper4, we developed our algorithm. To speed up the algorithm, we use parallel package to parallelize computations the factorization function.

### 3. Parameter Tuning
We tested the following parameters here.
f = 5, 10, 15
lambda_als = 1,5,10
lambda_p = 1, 5, 10
sigmas = 1,2,3
```{r, eval=FALSE}
fs <- c(5, 10, 15)
lambdas_als <- c(1, 5, 10)
lambdas_P <- c(1, 5, 10)
sigmas <- c(1, 2, 3)

cv.df <- expand.grid("fs"=fs, 
                     "lambdas_als"=lambdas_als, 
                     "lambdas_p"=lambdas_P, 
                     "sigmas"=sigmas)
```

The code of parameter tuning is
```{r, eval=FALSE}
cl <- makeCluster(cores)

x <- cv.df %>%
  as_tibble() %>%
  mutate(train_mean=NA,
         test_mean=NA)

for (i in 1:dim(x)[1]){
  tmp <- cv.functionYQ(dat_train=data,
                       K=5,
                       f=x[i,]$fs,
                       maxIter=15,
                       lambdas_als=x[i,]$lambdas_als,
                       lambdas_p=x[i,]$lambdas_p,
                       sigmas=x[i,]$sigmas)
  
  x[i,"train_mean"]=tmp$mean_train_rmse
  x[i,"test_mean"]=tmp$mean_test_rmse

  rm(tmp)
}
save(x, file="../output/cvtmp.RData")

stopCluster(cl)
```

The result is like
```{r}
load(file="../output/cvtmp.RData")
x
```

The optimal parameters for ALS + KRR is
```{r,warning=FALSE}
x[which.min(x$test_mean), ][1:4]
```

### 4. Optimal Model
 - Model Fitting
Here, we fit the ALS with post-processing, KRR
```{r, eval=FALSE, warning=FALSE}
cl <- makeCluster(cores)

alsP3Result <- ALS_KRR(data, train, test, f=15, maxIters=10, 
                       lambda_als=10, lambda_p=1, sigma=1)

stopCluster(cl)
save(alsP3Result, file = "../output/alsP3Result.RData")
```
 
 - Model Evaluation
 The final result of our model 1 is
```{r ALS with KRR result}
load(file="../output/alsP3Result.RData")
cat("RMSE for train set:", round(alsP3Result$train_RMSE, 4), 
    "  RMSE for test set:", round(alsP3Result$test_RMSE, 4))
```


#Model 2:Alternating Least Squares + Penalty of magnitudes + Temporal Dynamics + Kernel Ridge Regression

## Step 1. Load Data and Train-Test Split

## Step 2. Model Setup
### 1. Equations recap
### Alternating Least Squares

$$min_{q^{\star}p^{\star}} \sum_{(u,i) \in K} (r_ui-q_i^Tp_u)^2 + \lambda(\sum _i n_{q_i} \| q_i\|^2 +\sum _i n_{p_u} \| p_u\|^2 )$$
ALS technique rotate between fixing the $q_i$'s and fixing the $p_u$'s. When all pu's are fixed, system recomputes the $q_i$'s by solving a least-squares problem, and vice versa. This ensures that each step decreases object function until convergence. 

$f$: dimension of latent factors

$q_i$: factors associated with item i , measures the extent to which items possesses those factors

$p_u$: factors associated with user u, measures the extent of interest that user has in an item  are high on corresponding factors.


### Regularizations 
 - Penalty of magnitudes:

$$\sum_{(u,i)\in K} \lambda (\|q_i\|^2+\|p_u\|^2)$$ 
The constant lambda controls the extent of regularization and we determined it through cross validation. 

 - Temporal Dynamics

$$\hat{r}_{ui} = q_i^Tp_u + \mu + b_u + b_i + \alpha_udev_u(t)$$ 
The $dev_u(t)$ is the related to the difference between rating time of user and their average rating time, measured by 
$$dev_u(t) = sign(t - t_u)·|t-t_u|^\beta$$ 

The final objective function of our model 2 is
$$min_{q,p} \sum_{(u,i) \in K} [r_{ui}-(q_i^Tp_u + \mu + b_u + b_i + \alpha_udev_u(t)]^2 + \lambda\sum( \| q_i\|^2 +\| p_u\|^2 + b_i^2 + b_u^2 + \alpha_u^2)$$

### 2. Algorithm Setup
For matrix factorization, with the method in paper4, to include $b_u$, $b_i$, and $a_u$ in the iterations, we add one row to the movie matrix and two rows to the user matrix. To speed up the algorithm, we use parallel package to parallelize computations and tibble in the factorization function.

### 3. Parameter Tuning
To save the time, with the refult of parameters tuning of ALS and KRR from the former model and some attempts, we only tested the following parameters here.
lambda_als = 1,5,10
beta = 0.2,0.4,0.5
And set
f = 15
lambda_p = 1
sigmas = 1
```{r}
fs <- c(15)
lambdas_P <- c(1)
sigmas = c(1)


lambdas_als <- c(1, 5,10)
betas <- c(0.2,0.4,0.5)

cv.df <- expand.grid("fs"=fs, 
                     "lambdas_als"=lambdas_als, 
                     "lambdas_p"=lambdas_P, 
                     "sigmas" =sigmas,
                     "betas"=betas)

```

The code of parameter tuning is
```{r,eval=FALSE}
set.seed(1)
cl <- makeCluster(cores)

arp_result=cv.df %>%
  as_tibble() %>%
  mutate(train_mean=NA,test_mean=NA)

for (i in 1:dim(arp_result)[1]){

  tmp <- cv.functionA3R3P3(dat_train=data,
                       K=4,
                       f=x[i,]$fs,
                       maxIter=15,
                       lambdas_als=x[i,]$lambdas_als,
                       lambdas_p=x[i,]$lambdas_p,
                       sigmas=x[i,]$sigmas,
                       betas=x[i,]$betas)

  arp_result[i,"train_mean"]=tmp$mean_train_rmse
  arp_result[i,"test_mean"]=tmp$mean_test_rmse

  rm(tmp)
}

save(arp_result, file="../output/cvARP.RData")

stopCluster(cl)
```

The result is 
```{r}
load(file="../output/cvARP.RData")
arp_result
```

The optimal parameters for ALS + R1R3 + KRR is
```{r,warning=FALSE}
arp_result[which.min(arp_result$test_mean),][1:5]
```

### 4. Optimal Model
 - Model Fitting
Here, we fit the complete model with post-processing, KRR
```{r ALS with R1R3 With KRR, eval=FALSE}
cl <- makeCluster(cores)
ALS.R1R3.KRR.model <- ALS_R_KRR(data, data_train, data_test, f = 15,
                                maxIters = 10, lambda_als = 10, lambda_p = 1, 
                                sigma = 1, beta = 0.2)
stopCluster(cl)
save(ALS.R1R3.KRR.model, file="../output/ALS_R1R3_KRR_model.RData")

stopCluster(cl)
```
 - Model Evaluation
 We can see that when we add post-processing to the matrix factorization, the performance decrease.
 The final result of our model 2 is
```{r ALS with R1R3 and KRR result}
load(file="../output/ALS_R1R3_KRR_model.RData")
cat("RMSE for train set:", round(ALS.R1R3.KRR.model$train_RMSE, 4),
    "  RMSE for test set:", round(ALS.R1R3.KRR.model$test_RMSE, 4))
```

We noticed that without P3, the matrix factorization alone performs better. With our previous attempts, we chose the following parameters. (We also noticed that the optimal parameters for matrix factorization are different with or without the post-processing)
```{r R1R3 only, eval=FALSE}
set.seed(1)
cl <- makeCluster(cores)

ALS.R1R3.only <- ALS.R1R3(f = 10, lambda = 10, beta = 0.4, maxIters = 15 , data = data,
                          train = data_train, test = data_test)

stopCluster(cl)
save(ALS.R1R3.only, file="../output/ALS_R1R3_only.RData")
```

Visualization of training and test RMSE by different iterations
```{r ALS with R1R3 only plot}
load(file="../output/ALS_R1R3_only.RData")
g2 <- data.frame(iteration = c(1:15),Training_MSE = ALS.R1R3.only$`Train RMSE`,
                 Test_MSE = ALS.R1R3.only$`Test RMSE`) %>%
  gather(key = Train_Test,value = RMSE,-iteration) %>%
  ggplot(aes(x = iteration, y = RMSE, col = Train_Test)) + geom_point() +
  theme_light()
g2
```

We can see, without post-processing, the regularization can significantly reduce the RMSE in the second iteration. The final result is
```{r ALS with R1R3 only result}
load(file="../output/ALS_R1R3_only.RData")
cat("RMSE for train set:", round((ALS.R1R3.only$`Train RMSE`)[15], 4), 
    "  RMSE for test set:", round((ALS.R1R3.only$`Test RMSE`)[15], 4))
```

# Summary
|   | A3 + KRR | A3 + R1R3 + KRR| A3 + R1R3 |
| ------ | :------: | :------: | :------: | 
| training RMSE | 0.6216 | 0.5002 | 0.6509 |
| test RMSE | 0.9698 | 1.2418 | 0.851 |

By comparing RMSEs, we know that, given (A3+P3), adding (R1+R3) tends to overfit. With (R1+R3), the training RMSE is lower but the test RMSE is higher. But without KRR, (A3 + R1R3) performs better than (A3 + KRR). (A3 + R1R3) is the most robust model, with a higher training MSE but lower test RMSE. Thus, we think KRR and R1R3 may not work well together, but can improve the original matrix factorization separately. 
