---
title: "ALS+R1R3+KRR"
author: "Group 4"
output: pdf_document
---

```{r set up, include = F}
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(cache = T)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(ggplot2)
library(anytime)
library(caret)
library(parallel)
library(krr)
library(dplyr)

cl <- makeCluster(4)
```

#Model 2:Alternating Least Squares + Penalty of magnitudes + Temporal Dynamics + Kernel Ridge Regression

## Step 1. Load Data and Train-Test Split
In this step, we loaded the data and added variable timediff for temporal dynamics, which is the difference between the user's rating time and the average rating time of this user. Then we splitted the original dataset to training set(80%) and test set(20%)
```{r}
data <- read.csv("../data/ml-latest-small/ratings.csv")
# Convert timestamp and set up time difference
data$timestamp <- anydate(data$timestamp)
data <- data %>% 
  group_by(userId) %>% 
  mutate(timediff = (timestamp - mean(timestamp)) %>% as.numeric) %>% 
  ungroup() %>%
  arrange(timestamp)

set.seed(1)

# train-test split (.8/.2)
train_ind <- createDataPartition(data$userId, p=.8, list=F)


data_train <- data[train_ind, ]
data_test <- data[-train_ind, ]
```

## Step 2. Model Setup
### 1. Equations recap
### Alternating Least Squares

$$min_{q^{\star}p^{\star}} \sum_{(u,i) \in K} (r_ui-q_i^Tp_u)^2 + \lambda(\sum _i n_{q_i} \| q_i\|^2 +\sum _i n_{p_u} \| p_u\|^2 )$$
ALS technique rotate between fixing the $q_i$'s and fixing the $p_u$'s. When all pu's are fixed, system recomputes the $q_i$'s by solving a least-squares problem, and vice versa. This ensures that each step decreases object function until convergence. 

$f$: dimension of latent factors

$q_i$: factors associated with item i , measures the extent to which items possesses those factors

$p_u$: factors associated with user u, measures the extent of interest that user has in an item  are high on corresponding factors.


### Regularizations 
 - Penalty of magnitudes:

$$\sum_{(u,i)\in K} \lambda (\|q_i\|^2+\|p_u\|^2)$$ 
The constant lambda controls the extent of regularization and we determined it through cross validation. 

 - Temporal Dynamics

$$\hat{r}_{ui} = q_i^Tp_u + \mu + b_u + b_i + \alpha_udev_u(t)$$ 
The $dev_u(t)$ is the related to the difference between rating time of user and their average rating time, measured by 
$$dev_u(t) = sign(t - t_u)·|t-t_u|^\beta$$ 

The final objective function of our model 2 is
$$min_{q,p} \sum_{(u,i) \in K} [r_{ui}-(q_i^Tp_u + \mu + b_u + b_i + \alpha_udev_u(t)]^2 + \lambda\sum( \| q_i\|^2 +\| p_u\|^2 + b_i^2 + b_u^2 + \alpha_u^2)$$


### Post-processing
Kernel Ridge Regression:
$$\hat{\beta} = (X^{T}X + \lambda I)^{-1}X^{T}y $$
$$\hat{y_{i}} = {x_i}^T\hat{\beta}$$
Equilvalent to:
$$\hat{\beta} = X^T(XX^T + \lambda I)^{-1}y$$ 

$$\hat{y}_{i} = K(x_i^{T}, X)(K(X,X) + \lambda I)^{-1}y$$ 
$y$: vector of movies rated by user i
$X$: a matrix of observations - each row of X is normalized vector of features of one movie j rated by user i: 
$$x_{j_2} = \frac{q_j}{\|q_j\|}$$ 

$K$: 
$$K(x_{i}^T, X) = x_i^Tx_j$$ 
$$K(x_{i}^T, X) = \exp(2(x_i^Tx_j -1))$$
$\lambda$:
$$\lambda = 0.5$$

### 2. Algorithm Setup
For matrix factorization, with the method in paper4, to include $b_u$, $b_i$, and $a_u$ in the iterations, we add one row to the movie matrix and two rows to the user matrix. To speed up the algorithm, we use parallel package to parallelize computations and tibble in the factorization function.

### 3. Parameter Tuning
To save the time, with the refult of parameters tuning of ALS and KRR from the former model and some attempts, we only tested the following parameters here.
lambda_als = 1,5,10
beta = 0.2,0.4,0.5
And set
f = 15
lambda_p = 1
sigmas = 1
```{r}
fs <- c(15)
lambdas_P <- c(1)
sigmas = c(1)


lambdas_als <- c(1, 5,10)
betas <- c(0.2,0.4,0.5)

cv.df <- expand.grid("fs"=fs, 
                     "lambdas_als"=lambdas_als, 
                     "lambdas_p"=lambdas_P, 
                     "sigmas" =sigmas,
                     "betas"=betas)

```
The code of parameter tuning
```{r}
source("../lib/cross_validation.R")
source("../lib/ALS_R_KRR.R")
source("../lib/ALSwithR1R3.R")
source("../lib/ALSwithP3.R")

set.seed(1)
# cl <- makeCluster(4)
# 
# arp_result=cv.df %>%
#   as_tibble() %>%
#   mutate(train_mean=NA,test_mean=NA)
# 
# for (i in 1:dim(arp_result)[1]){
# 
#   tmp <- cv.functionA3R3P3(dat_train=data,
#                        K=4,
#                        f=x[i,]$fs,
#                        maxIter=15,
#                        lambdas_als=x[i,]$lambdas_als,
#                        lambdas_p=x[i,]$lambdas_p,
#                        sigmas=x[i,]$sigmas,
#                        betas=x[i,]$betas)
#   
#   arp_result[i,"train_mean"]=tmp$mean_train_rmse
#   arp_result[i,"test_mean"]=tmp$mean_test_rmse
# 
#   rm(tmp)
# }
# 
# save(arp_result, file="../output/cvARP.RData")
# 
# stopCluster(cl)

load(file="../output/cvARP.RData")
arp_result
```
The optimal parameters for ALS + R1R3 + KRR is
```{r,warning=FALSE}
arp_result[which.min(arp_result$test_mean),]
```

### 4. Optimal Model
 - Model Fitting
 To see the how ALS with R1R3 works, we first conducted the matrix factorization separately
```{r ALS with R1R3, warning=FALSE}
# ALS with R1R3
# cl = makeCluster(8)
# ALS.R1R3.model <- ALS.R1R3(f = 15, lambda = 1, beta = 0.2, maxIters = 15 , data = data,train = data_train, test = data_test)
# stopCluster(cl)
# save(ALS.R1R3.model, file="../output/ALS_R1R3_model.RData")
# stopCluster(cl)

load(file="../output/ALS_R1R3_model.RData")
```
 Visualization of training and test RMSE by different iterations
```{r ALS with R1R3 plot}
g1 <- data.frame(iteration = c(1:15),Training_MSE = ALS.R1R3.model$`Train RMSE`,
                 Test_MSE = ALS.R1R3.model$`Test RMSE`) %>%
  gather(key = Train_Test,value = RMSE,-iteration) %>%
  ggplot(aes(x = iteration, y = RMSE, col = Train_Test)) + geom_point() +
  theme_light()
g1
```
Here, we fit the complete model wiht post-processing, KRR
```{r ALS with R1R3 With KRR, warning=FALSE,message=FALSE,results = 'hide'}
# cl = makeCluster(8)
# ALS.R1R3.KRR.model <- ALS_R_KRR(data, data_train, data_test, f = 15, maxIters = 15, lambda_als = 1, lambda_p = 1, sigma = 1,beta = 0.2)
# stopCluster(cl)
# save(ALS.R1R3.KRR.model, file="../output/ALS_R1R3_KRR_model.RData")
# 
# stopCluster(cl)

load(file="../output/ALS_R1R3_KRR_model.RData")
```
 - Model Evaluation
 We can see that when we add post-processing to the matrix factorization, the performance decrease.
 The final result of our model 2 is
```{r ALS with R1R3 and KRR result}
cat("RMSE for train set:", round(ALS.R1R3.KRR.model$train_RMSE, 4), "  RMSE for test set:", round(ALS.R1R3.KRR.model$test_RMSE, 4))
```
 
 


